### 2020-04-30 serial.tex
still working on discussion

Characterizing generation- and serial-interval distributions is critical to understanding the outset of an outbreak as they determine the time scale of disease transmission.
Here, we show that the initial forward serial-interval distributions --- measured during exponential growth phase of an epidemic --- provides the correct link between the exponential growth rate $r$ and the reproduction number \RR.

Our study demonstrates the importance of defining a reference cohort for studying any epidemiological time delay distributions.
Previous studies have shown that generation interval can be either measured forward or backward \citep{nishiura2010time,champredon2015intrinsic,britton2019estimation};
we generalize their ideas by showing that these ideas can be applied to all distributions.
Changes in the backward delay distributions is particularly robust across all distributions because they depend on previous changes in cohort sizes, which, in turn, depend on incidence of infection.
Previous analyses of COVID-19 epidemics have attempted to reconstruct the time series of symptomatic cases or infected cases from the reports of confirmed cases by assuming a constant backward delay distribution (e.g., \cite{tempvar, park2020potential, shim2020transmission}).
Although such approaches may be able to roughly match the time scale of an epidemic, we recommend using deconvolution approaches instead for proper reconstruction of epidemic time series \citep{goldstein2009reconstructing}.

Using cohort-based approaches, we define forward and backward serial intervals.
The forward serial-interval distribution describes the renewal process of the symptomatic individuals, and therefore, its initial distribution  provides the correct link between $r$ and \RR.
While our results support the use of serial interval distributions for calculating \RR, 
they also reveal gaps in current approaches to characterizing serial intervals.
Previous studies have (i) recommended using up-to-date serial intervals \citep{thompson2019improved}, (ii) assumed that the serial interval distribution remains constant throughout an epidemic \citep{wallinga2004different, cori2013new}, and (iii) applied serial interval distributions estimated from one location to analyzing epidemics in other locations \citep{tempvar};
however, all of these approaches implicitly ignore spatiotemporal variation in incidence, which affects the serial-interval distributions, and may therefore give systematically biased estimates of \RR. 
For example, given large geographical differences in the observed exponential growth rates of the COVID-19 epidemic \citep{tempvar}, the initial forward serial interval distributions is likely to vary across different countries.
Future studies should aim to characterize spatiotemporal variation in serial intervals and understand the drivers of their differences.

>>>>

Characterizing generation- and serial-interval distributions is critical to understanding the course of an outbreak as they determine the time scale of disease transmission.
Here, we show that the initial forward serial-interval distributions --- measured during exponential growth phase of an epidemic --- provides the correct link between the exponential growth rate $r$ and the reproduction number \RR.
We further illustrate that epidemiological dynamics can cause the mean backward incubation period to increase over time, which in turn decreases the mean forward serial interval.
Failing to account for these changes can result in underestimation of \RR.

Our study demonstrates the importance of defining a reference cohort for characterizing any epidemiological time distributions.
Previous studies have shown that generation interval can be either measured forward or backward \citep{nishiura2010time,champredon2015intrinsic,britton2019estimation};
we generalize their ideas and show that they can be applied to any distributions.
Changes in the backward delay distributions is expected to be a robust feature across any outbreaks because they are primarily driven by changes in cohort sizes, which, in turn, depend on incidence of infection.
Recent analyses of COVID-19 epidemics have attempted to reconstruct epidemic time series of symptomatic cases or infected cases from the reports of confirmed cases by using time-independent delay distributions from infection or symptom onset to reporting (e.g., \cite{tempvar, park2020potential, shim2020transmission});
these studies effectively assume constant backward delay distributions.
Although such approaches may be able to roughly match the time scale of an epidemic, we recommend using deconvolution approaches \citep{goldstein2009reconstructing} or mechanstic modeling approaches for proper reconstruction of epidemic time series instead \citep{flaxman2020estimating}.

Using cohort-based approaches, we define forward and backward serial intervals.
The forward serial-interval distribution describes the renewal process between symptomatic cases, and therefore, its initial distribution provides the correct link between $r$ and \RR.
While our results support the use of serial interval distributions for calculating \RR, 
they also reveal gaps in current practices in incorporating serial-interval distributions into outbreak analyses.
In particular, \cite{thompson2019improved} recently emphasized the importance of using up-to-date serial interval data for accurate estimation of time-varying reproduction numbers;
however, our study demonstrates that early serial interval data is required to estimate \RR and using up-to-date serial interval data can, in fact, underestaimte \RR.
Moreover, \cite{tempvar} recently analyzed case reports of COVID-19 from over 30 countries to estimate their time-varying reproduction numbers but relied on a serial-interval distribution;
given large geographical differences in the observed exponential growth rates of the COVID-19 epidemic, the initial forward serial interval distributions is likely to vary across different countries.

Future studies should aim to characterize spatiotemporal variation in serial intervals and understand the drivers of their differences.
